Molecular dynamics simulation of synchronization in driven particles

### 2021-02-13 manuscript/manuscript_ajp_2020.tex
starting to fill in phase section

In Section ~\ref{sec:problems}
we explore our numerical model.
Analytic exercises examine the 
to linear drag equation in Ex.~\ref{ex:reynolds}, 
the equation of motion in Ex.~\ref{ex:n2l},
and 
numerical integration techniques in Ex.~\ref{ex:euler}.
Numerical exercises extend the model
to include finite temperature effects
in Ex.~\ref{ex:brownian}.
We explore changes in the particle motion
with parameter changes in Ex.~\ref{ex:parameters}.
Methods to visualize the dynamics of the particle motion
are in Ex.~\ref{ex:animation}.

@@

%\begin{enumerate}
%\item
\subsection{Drag Models and Reynolds Numbers.}
  Stokes' law describes the viscous drag on a sphere
  moving at velocity $\vec{v}$ as 
  $\vec{F}^{lin} = -3 \pi \eta D \vec{v}$ 
  where $\eta$ is the dynamic fluid viscosity and 
  $D$ is the particle diameter \cite{Taylor2005}.
  In simulation we
  subsume the constants $3 \pi D$
  such that $3 \pi D \eta \rightarrow \eta $.
  %
  Often drag forces are
  modeled as a polynomial series ~\cite{Taylor2005}
  \begin{equation}
    \vec{F}^{drag} = -b \vec{v} - c v^2 \hat{v} + \ldots  .
  \end{equation}
  %This model should be familiar to readers
  %of a standard classical mechanics text, such as Ref.
  %when studying 
  %models Millikan's oil drops.
  Truncating the series to the first term
  is justified by demonstrating the sphere
  has a low Reynolds number  
  $R = D v \rho / \eta$
  where $\rho$ is the fluid density and $v$ the particle's speed.
  When $R$ is small, the quadratic and higher order terms
  may be ignored in favor of the linear drag term.

@@

  Synchronization 
  %is often studied with simple computational models and
  can be detected with 
  with phase
  plots or Lissajous figures.
  %First reported in 1815 by Nathanial Bowditch
  %and later by Lissajous in 1857  \cite{Lissajous1857}.
  Lissajous figures
  combine two periodic signals
  into a two dimensional plot.  
  %and easily generated with a computer code 
  %or oscilloscope \cite{Tong1997}. 
  A mode-locked system is confined to 
  a closed loop in  
  parametric phase space % showing the relationship
  %between two periodic functions
  with patterns determined by the mode. % frequency ratio,
  %with a frequency generator,
  %where two combined frequencies at perpendicular orientations
  Note the term phase
  has two scientific meanings:  
  %``oscillation phase'' and ``phase space.''
  %refer to different uses of the word ``phase.''
  oscillation phase $\phi_0$ is the constant of the argument
  in a periodic function
  $\sin{(\omega t + \phi_0)}$ and phase space is
  used to characterize oscillating systems with
  parametric plots.

  In a non-driven system, an appropriate
  phase space is $v_y$ versus $y$
  in order to characterize whether a particle velocity
  synchronizes to a particular spatial location.
  In a system driven parallel to the $y-$direction,
  the particle position is monotonically increasing
  in time.
  A synchronized driven particle
  repeats a particular velocity
  at the same $y-$position of the particle
  within the substrate period.
  We define a phase variable 
  \begin{equation}
    %(y mod \lambda) / \lambda
    \phi(t) = 2\pi(y(t)-\bar{v}_y t)/\lambda
  \end{equation}
  that is centered about the average particle velocity $\bar{v}_y$
  and normalized by the substrate period $\lambda$ \cite{Juniper2015}.
  We plot this phase
  versus the phase velocity 
  \begin{equation}
    %(y mod \lambda) / \lambda
    \dot{\phi}(t) = 2\pi(v_y(t)-\bar{v}_y/\lambda.
  \end{equation}
  %Juniper %REF 53

>>>>

In Section ~\ref{sec:problems}
we explore our model
with suggested exercises for interested readers.
We include analytic exercises to examine the 
to linear drag equation in Ex.~\ref{ex:reynolds}, 
the equation of motion in Ex.~\ref{ex:n2l},
and 
numerical integration techniques in Ex.~\ref{ex:euler}.
We extend the model with
numerical exercises %extend the model
to include finite temperature effects
in Ex.~\ref{ex:brownian}.
We explore changes in the particle motion
with parameter changes in Ex.~\ref{ex:parameters}.
Methods to visualize the dynamics of the particle motion
are in Ex.~\ref{ex:animation}.

@@

%\begin{enumerate}
%\item
\subsection{Drag Models and Reynolds Numbers.}
Stokes' law describes the drag force
$\vec{F}^{lin} = -3 \pi \eta D \vec{v}$ 
on a sphere
moving through a viscous liquid at velocity $\vec{v}$ 
  where $\eta$ is the dynamic fluid viscosity and 
  $D$ is the particle diameter \cite{Taylor2005}.
  In simulation we
  subsume the constants $3 \pi D$
  such that $3 \pi D \eta \rightarrow \eta $.
  %
  Often drag forces are
  modeled as a polynomial series ~\cite{Taylor2005}
  \begin{equation}
    \vec{F}^{drag} = -b \vec{v} - c v^2 \hat{v} + \ldots  .
  \end{equation}
  %This model should be familiar to readers
  %of a standard classical mechanics text, such as Ref.
  %when studying 
  %models Millikan's oil drops.
  Truncating the series to the first term
  is justified by demonstrating the sphere
  has a low Reynolds number  
  $R = D v \rho / \eta$
  where $\rho$ is the fluid density and $v$ the particle's speed.
  When $R$ is small, the quadratic and higher order terms
  may be ignored in favor of the linear drag term.

@@

  Synchronization 
  can be detected with 
  phase
  plots or Lissajous figures
  that 
  %First reported in 1815 by Nathanial Bowditch
  %and later by Lissajous in 1857  \cite{Lissajous1857}.
  %Lissajous figures
  compare two periodic signals
  into a two dimensional plot.
  In a non-driven system, an appropriate
  phase space is $v_y$ versus $y$.  
  %and easily generated with a computer code 
  %or oscilloscope \cite{Tong1997}.
  When a particle velocity
  synchronizes to a particular spatial location
  the dynamics
  appears as 
  %A mode-locked system is confined to 
  a closed loop in  
  phase space, % showing the relationship
  %between two periodic functions
  with patterns determined by the mode. % frequency ratio,
  %with a frequency generator,
  %where two combined frequencies at perpendicular orientations
  %\footnote{The term phase
  %has two scientific meanings:  
  %oscillation phase $\phi_0$ is the constant of the argument
  %in a periodic function
  %$\sin{(\omega t + \phi_0)}$ and phase space is
  %used to characterize oscillating systems with
  %parametric plots.}.
  In a system driven parallel to the $y-$direction,
  the particle position is monotonically increasing
  in time.
  A synchronized driven particle
  repeats a particular velocity
  at the same $y-$position of the particle
  within the substrate period.
  We define a phase variable 
  \begin{equation}
    %(y mod \lambda) / \lambda
    \phi(t) = 2\pi(y(t)-\bar{v}_y t)/\lambda
  \end{equation}
  that is centered about the average particle velocity $\bar{v}_y$
  and normalized by the substrate period $\lambda$ \cite{Juniper2015}.
  We plot this phase
  versus the phase velocity 
  \begin{equation}
    %(y mod \lambda) / \lambda
    \dot{\phi}(t) = 2\pi(v_y(t)-\bar{v}_y/\lambda.
  \end{equation}
  %Juniper %REF 53
